Variation of Euclidean Algorithm for finding gcd of two numbers

### 2017-12-14 latex/draft.tex
define mod dash

@@

Let's look at a more detailed algorithm, iterative in nature and presented in a book by M. Dietzfelbinger, and is based on the previously mentioned Euclid's algorithm which is recursive in nature but the principle remains same.
\newline
\newline \textbf{Algorithm 2:}
\newline \textbf{GCD Algorithm (Euclidean Algorithm)}
\newline INPUT: Two integers \var{n}, \var{m}
\newline METHOD:

@@

For \textbf{t} = 7, b\textsubscript{t} = 0 and the algorithm exits with the return value 
46 as the gcd for (12742, 10534). The captured output sequences are presented in the table 2 as below.
% ---------------------------------TABBLE 2 -----------------------------------
% [t] indicates the top of the page
% {table*} used for wide table
\begin{table}[!htbp]
\caption{Table showing intermediate values while calculating gcd for (10534, 12742)}
\label{table_org_book_example}
\centering
	\begin{tabular}{c | c | c}
		\hline
		$i$ & $a\textsubscript{i}$ & $b\textsubscript{i}$ \\ [0.5ex] 
		\hline
		0 & 12742 & 10534 \\ 
		1 & 10534 & 2208 \\  
		2 & 2208  & 1702 \\
		3 & 1702  & 506 \\
		4 & 506   & 184 \\
		5 & 184   & 138 \\
		6 & 138   & 46 \\
		7 & 46    & 0
	\end{tabular}
\end{table}
% ---------------------------------TABBLE 2 -----------------------------------

Now it is the time to see the results when the claim made in this paper is being applied 
to the algorithm. According to the claim, instead of the remainder of the division, the seek value (\var{s}) that could make the dividend a multiple of divisor, which is same as difference between divisor and remainder, can be used in the algorithm without affecting the result. So we can modify line no. 8 of the Algorithm 2 by using a modified implementation of \mod and we call it \mod' and can be expressed as below. Let's name this modified algorithm as algorithm no: 3 .  \linebreak
\textit{a} mod' \textit{b} = \textit{s}, where  \textit{s} = \textit{b - r}, \textit{r} = \textit{a} mod \textit{b}

@@

\textbf{Alternative Approach using 's'}
Now the question raises, what is the significance of this \var{s} over \var{r}.
If we analyse the Euclidean algorithm for finding greatest common divisor, it is quite evident that, the arguments are getting reduced and one of them reduced till 0 when finally the output of the algorithm is obtained. So it is quite natural that if \var{s} can be substituted  for \var{r} then finding the smallest of them at each recursive call and using it for the subsequent recursion can help the recursion bringing the operand closer to 0 more quickly. The modified algorithm can be presented below.
\newline \textbf{Algorithm 4:}
\begin{algorithmic}[1]
\STATE \progvar{a}, \progvar{b}: integer
\IF{$|n| \geq |m|$}
	\STATE $\progvar{a} \gets |n|; \progvar{b} \gets |m|;$
\ELSE
	\STATE $\progvar{a} \gets |m|; \progvar{b} \gets |n|;$
\ENDIF
\WHILE{$\progvar{b} > 0$ } 
	\STATE $(\progvar{a}, \progvar{b}) \leftarrow (\progvar{b}, \progvar{a}  $ $mod'' $ $\progvar{b});$
\ENDWHILE
\RETURN \progvar{a}
\end {algorithmic}

So it can be noted from the above that there is apparent no change from the previously given algorithm except line no. 8. The modified algorithm uses $mod'$ instead of \mod for the modulus operation. This $mod''$ can be expressed as below.
If \textit{r} = \textit{a} mod \textit{b}, \textit{s} = \textit{b - r};
\textit{a} $mod''$ \textit{b} = min(r, s), where \textit{min} is the function returns the least of the arguments.

>>>>

\newcommand{\moddash}{\text{mod'}}

\newcommand{\moddashdash}{\text{mod''}}

@@

Let's look at a more detailed algorithm, iterative in nature and presented in a book [2] by M. Dietzfelbinger, and is based on the previously mentioned Euclid's algorithm which is recursive in nature but the principle remains same.
\newline
\newline \textbf{Algorithm 2:}
\newline \textbf{GCD Algorithm (Euclidean Algorithm)}
\newline INPUT: Two integers \var{n}, \var{m}
\newline METHOD:

@@

For \textbf{t} = 7, b\textsubscript{t} = 0 and the algorithm exits with the return value 
46 as the gcd for (12742, 10534). The captured output sequences are presented in the table 2 (referred from book[2]) as below.
% ---------------------------------TABBLE 2 -----------------------------------
% [t] indicates the top of the page
% {table*} used for wide table
\begin{table}[!htbp]
\caption{Table showing intermediate values while calculating gcd for (10534, 12742)}
\label{table_org_book_example}
\centering
	\begin{tabular}{c | c | c}
		\hline
		$i$ & $a\textsubscript{i}$ & $b\textsubscript{i}$ \\ [0.5ex] 
		\hline
		0 & 12742 & 10534 \\ 
		1 & 10534 & 2208 \\  
		2 & 2208  & 1702 \\
		3 & 1702  & 506 \\
		4 & 506   & 184 \\
		5 & 184   & 138 \\
		6 & 138   & 46 \\
		7 & 46    & 0
	\end{tabular}
\end{table}
% ---------------------------------TABBLE 2 -----------------------------------

Now it is the time to see the results when the claim made in this paper is being applied 
to the algorithm. According to the claim, instead of the remainder of the division, the seek value (\var{s}) that could make the dividend a multiple of divisor, which is same as difference between divisor and remainder, can be used in the algorithm without affecting the result. 
For this variation, let's define a new operator slightly different from \mod, and we name this as \moddash(say \mod -dash).
\begin{equation} \label{eq:gcd_modulus_1_1}
\var{a} \> \moddash \> \var{b} = \var{s}, \>  {s} = \var{b} - \var{r}, \var{r} = \var{a} \> \mod \> \var{b}
\end{equation}

So we can modify line no. 8 of the Algorithm 2 by using \moddash. Let's name this modified algorithm as algorithm no: 3 .  \linebreak

@@

\textbf{Alternative Approach using 's'}
Now the question raises, what is the significance of this \var{s} over \var{r}.
If we analyse the Euclidean algorithm for finding greatest common divisor, it is quite evident that, the arguments are getting reduced and one of them reduced till 0 when finally the output of the algorithm is obtained. So it is quite natural that if \var{s} can be substituted  for \var{r} then finding the smallest of them at each recursive call and using it for the subsequent recursion can help the recursion bringing the operand closer to 0 more quickly.

Let's modify the algorithm, with a small variation with a slightly modified representaion of \mod and \moddash; let's define a new operator as \moddashdash(say \mod -dash-dash).

\begin{equation} \label{eq:gcd_modulus_1_2}
\var{a} \> \moddashdash  \> \var{b} = \text{min} \var(r, s);
\var{r} = \var{a} \> \mod \> \var{b}; \> \var{s} = \var{b - r}
\end{equation}

where \textit{min} is the function returns the least of the arguments.

The modified algorithm can be presented below.
\newline \textbf{Algorithm 4:}
\begin{algorithmic}[1]
\STATE \progvar{a}, \progvar{b}: integer
\IF{$|n| \geq |m|$}
	\STATE $\progvar{a} \gets |n|; \progvar{b} \gets |m|;$
\ELSE
	\STATE $\progvar{a} \gets |m|; \progvar{b} \gets |n|;$
\ENDIF
\WHILE{$\progvar{b} > 0$ } 
	\STATE $(\progvar{a}, \progvar{b}) \leftarrow (\progvar{b}, \progvar{a}  $ $mod'' $ $\progvar{b});$
\ENDWHILE
\RETURN \progvar{a}
\end {algorithmic}

So it can be noted from the above that there is apparent no change from the previously given algorithm except line no. 8. The modified algorithm uses $mod'$ instead of \mod for the modulus operation. This $mod''$ can be expressed as below.
